---
title: "510 Project (Worker Productivity)"
output: html_document
date: "2022-12-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro to data set

Source:
https://www.kaggle.com/datasets/ishadss/productivity-prediction-of-garment-employees

Variables:

01 date : Date in MM-DD-YYYY

02 day : Day of the Week

03 quarter : A portion of the month. A month was divided into four quarters

04 department : Associated department with the instance

05 teamno : Associated team number with the instance 

06 noofworkers : Number of workers in each team 

07 noofstylechange : Number of changes in the style of a particular product

08 targetedproductivity : Targeted productivity set by the Authority for each team for each day. 

09 smv : Standard Minute Value, it is the allocated time for a task 

10 wip : Work in progress. Includes the number of unfinished items for products 

11 overtime : Represents the amount of overtime by each team in minutes

12 incentive : Represents the amount of financial incentive (in BDT) that enables or motivates a particular course of action.

13 idletime : The amount of time when the production was interrupted due to several reasons 

14 idlemen : The number of workers who were idle due to production interruption

15 actual_productivity : The actual % of productivity that was delivered by the workers. It ranges from 0-1.


# Some Questions. 

1. Which department is most productive?
2. Which day of the week is most productive for each department? 
3. Which team in the sewing department is the most productive?
4. Does the amount of over time impact the productivity in the sewing department? 
5. Does the amount of incentives impact the productivity in the sewing department? 
6. Does the number of workers impact the productivity in the sewing department? 

# Data cleaning

### 1. Read in the data set. Check if it's a tibble. 
```{r, message=FALSE, warning=FALSE}
library(readr)
work <- read_csv("~/Desktop/STAT510_F22/datasets/garments_worker_productivity.csv")
head(work)
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
is_tibble(work)
```


### 2. skim the data set. 

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
skim_without_charts(work)
```

There are a lot of missing values for wip. 

### 3. Investigate the missing values in wip

```{r, message=FALSE, warning=FALSE}
work |>
  mutate(is.na.wip = is.na(wip)) |>
  select(department, is.na.wip) |>
  group_by(department, is.na.wip) |>
  summarize(count = n())
```

The finishing department always has missing wip values because theytake all the "work in progress" parts and put them together. The sewing department will always have varying # of wip garments. 


### 4. select important variables and data cleaning

```{r}
work2 <- work |>
  select(-quarter, -smv, -idle_time, -idle_men, -no_of_style_change) |>
  mutate(department = ifelse(department == "sweing", "sewing", department))
names(work2)
```

```{r}
work3 <- work2 |>
  rename(dept = department, target = targeted_productivity, ot = over_time, 
         workers = no_of_workers, actual = actual_productivity) |>
  mutate(dept = factor(dept), 
         team = factor(team), 
         day = factor(day, 
                      levels = c("Saturday", "Sunday", "Monday", "Tuesday", 
                                 "Wednesday", "Thursday", "Friday"))) |>
  group_by(dept)
head(work3)
```

# Visualize distribution of variables of interest

```{r}
work3 |>
  ggplot( aes(x = dept)) +
  geom_bar()
```

sewing department has higher frequency 

```{r}
work3 |>
  ggplot( aes(x = day)) +
  geom_bar()
```



```{r}
work3 |>
  ggplot(aes(x = team)) + 
  geom_bar() + facet_wrap(~dept)
```

finishing department has more variation in observations per team



```{r}
work3 |>
  ggplot(aes(x = target)) + 
  geom_bar() + facet_wrap(~dept, scales = "free_x")
```

right skew in both departments. Most common target is 80% 



```{r}
work3 |>
  ggplot(aes(x = wip)) + 
  geom_histogram(bins = 100, na.rm = TRUE) + facet_wrap(~dept)
```

only sewing department has wip. extreme values are present.



```{r}
work3 |>
  ggplot(aes(x = ot)) + 
  geom_histogram(bins = 20) + facet_wrap(~dept, scales = "free_x")
```

left skew in finishing dept and somewhat normal distribution in sewing dept. Both dept's have extreme values



```{r}
work3 |>
  ggplot(aes(x = incentive)) + 
  geom_histogram(bins = 10) + facet_wrap(~dept, scales = "free_x")
```

incentive for finishing dept is mostly 0 with some extreme values
incentive for sewing dept seems to have a normal distribution



```{r}
work3 |>
  ggplot(aes(x = workers)) + 
  geom_histogram(bins = 30) + facet_wrap(~dept, scales = "free_x")
```

right skew and extreme value in sewing dept 



```{r}
work3 |>
 # filter(dept == "finishing") |>
  ggplot(aes(x = actual)) + 
  geom_histogram(bins = 20) + facet_wrap(~dept)

```

right skew in both dept 



# Visual Analysis of Questions

### 1.Which department is most productive? 

```{r}
work3 |>
  ggplot(aes(x = dept, y = actual)) + 
  geom_boxplot() + 
  stat_summary(fun.y="mean", shape = 1, color = "red")
```

The finishing department has higher productivity and no outliers. The sewing department might be more affected by changing conditions. 

### 2. Which day of the week is most productive for each department? 

```{r}
work3 |>
  ggplot(aes(x = day, y = actual)) +
  geom_boxplot() + facet_wrap(~dept) +
    stat_summary(fun.y="mean", shape = 1, color = "red")

```
For both departments, the median productivity is highest on Saturdays. This company seems to start its work week on Saturdays, because there is no data for Fridays in the data set. Perhaps the high productivity on Saturdays is related to it being right after a day off. 

There are a lot of outliers in the sewing department but only 1 in the finishing department

### 3. Which team in the sewing department is the most productive?

```{r}
library(ggplot2)
work.summary <- work3 |>
  group_by(dept, team) |>
  summarise(avg_actual= mean(actual)) |>
  ungroup()

work.summary

colors <- c("lightpink1", "lightpink2", "lightpink3", "mistyrose1", "mistyrose2", "mistyrose3", "rosybrown1", "rosybrown2", "rosybrown3", "thistle1", "thistle2", "thistle3")

work.summary |>
  ggplot(aes(x = team, y = avg_actual)) +
  geom_col(aes(fill = team)) +
  scale_fill_manual(values = colors) +
  facet_wrap(~dept, nrow = 2) + 
  theme_minimal() + theme(legend.position = "none") +
  ggtitle("Actual Productivity Per Department And Team")

```

```{r}
work.summary |>
  filter(dept == "sewing") |>
  mutate(team = reorder(team, desc(avg_actual))) |>
  ggplot(aes(x = team, y = avg_actual)) +
  geom_col(aes(fill = team)) +
  scale_fill_manual(values = colors) +
  theme_minimal() + theme(legend.position = "none") +
  ggtitle("Actual Productivity Per Team In Sewing Dept")
```

Team 1 is the most productive in the sewing department. 


# Modeling

### scatterplot matrix

```{r, message=FALSE, warning=FALSE}
library(GGally)
sewing <- work3 |>
  filter(dept == "sewing") |>
  ungroup() |>
  select(-dept, -date)

sewing |>
  ggpairs()
```

incentive looks linearly related to actual, but they are also highly correlated.

There are some 0 values in the incentive variable. We'd like to exclude rows where there was no incentive given to the workers, to evaluate the response with an actual incentive implemented by management. 

### exclude 0 incentive

```{r, message=FALSE, warning=FALSE}

sewing |>
  filter(incentive == 0) |>
  summarize(count = n())

sewing2 <- sewing |>
  filter(incentive != 0)

ggpairs(sewing2)
```

### AIC Variable selection

```{r}
n = nrow(sewing2)
mod0 = lm(actual ~ 1, data = sewing2)
mod.all = lm(actual ~., data = sewing2)
step(mod0, scope = list(lower = mod0, upper = mod.all))
```

```{r}
model.aic <- lm(actual ~ incentive + target + ot + team + workers, data = sewing2)
summary(model.aic)
```

### Check interaction

```{r}
add1(model.aic, ~.+incentive*target+incentive*ot+incentive*team+incentive*workers, test = 'F')

model.2 = update(model.aic, ~.+incentive:target+incentive:ot+incentive:workers)
summary(model.2)
```

```{r}
model.3 = update(model.2, ~.-incentive:ot)
summary(model.3)
```

### Best subset regression

```{r}
xmat = sewing2 |>
  select(-actual) |>
  select_if(is.numeric)
head(xmat)
```



```{r}
Team2 = ifelse(sewing2$team == 2, 1, 0)
Team3 = ifelse(sewing2$team == 3, 1, 0)
Team4 = ifelse(sewing2$team == 4, 1, 0)
Team5 = ifelse(sewing2$team == 5, 1, 0)
Team6 = ifelse(sewing2$team == 6, 1, 0)
Team7 = ifelse(sewing2$team == 7, 1, 0)
Team8 = ifelse(sewing2$team == 8, 1, 0)
Team9 = ifelse(sewing2$team == 9, 1, 0)
Team10 = ifelse(sewing2$team == 10, 1, 0)
Team11 = ifelse(sewing2$team == 11, 1, 0)
Team12 = ifelse(sewing2$team == 12, 1, 0)


xmat = xmat |>
  bind_cols(Team2, Team3, Team4, Team5, Team6, Team7, Team8, Team9, Team10, Team11, Team12)
head(xmat)

dim(xmat)
```

```{r}
library(leaps)
mod = regsubsets(xmat, sewing2$actual, nvmax = 16)
summary.mod = summary(mod)
names(summary.mod)

summary.mod

summary.mod$which
```

```{r}
summary.mod$rsq #check R^2
```

```{r}
summary.mod$adjr2 #check adjusted R^2
```

```{r}
plot(summary.mod$adjr2)


```

```{r}
plot(summary.mod$cp) #check Mallows' Cp
abline(1,1)
```




# Check LINE assumptions 

```{r}
library(broom)
model.table = augment(model.3)
names(model.table)

ggplot(model.table, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, colour = 'blue') +
  labs(x = 'Fitted Values', y = 'Residuals') +
  ggtitle('Residual vs Fit') +
  theme_bw()

```



```{r}
ggplot(model.table, aes(sample = .resid)) +
  stat_qq() +
  stat_qq_line() +
  ggtitle('Normal Q-Q Plot') +
  theme_bw()
```


```{r}
shapiro.test(resid(model.3))
```

# Transformation

```{r}
library(car)
mod.boxcox = boxCox(model.3, lambda = seq(-3, 3.5, length=20))

```

```{r}
lambda.opt = mod.boxcox$x[which.max(mod.boxcox$y)] 
lambda.opt
```



```{r}
sewing3 <- sewing2 |>
  mutate(actual.transf = actual ^ lambda.opt,
         incentive.ln = log(incentive),
         target.ln = log(target),
         ot.ln = log(ot), 
         workers.ln = log(workers))

sewing3

model.4 = lm(formula = actual.transf ~ incentive.ln + target + ot + team + workers.ln + 
    incentive:target + incentive:workers, data = sewing3)

summary(model.4)

```


# Check LINE assumptions on transformed model

```{r}
model.table.transf = augment(model.4)
names(model.table.transf)

ggplot(model.table.transf, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, colour = 'blue') +
  labs(x = 'Fitted Values', y = 'Residuals') +
  ggtitle('Residual vs Fit') +
  theme_bw()

```



```{r}
ggplot(model.table.transf, aes(sample = .resid)) +
  stat_qq() +
  stat_qq_line() +
  ggtitle('Normal Q-Q Plot') +
  theme_bw()
```


```{r}
shapiro.test(resid(model.transformed))
```


# Try filtering for a certain date range for sewing dept while excluding 0 incentive

```{r, message=FALSE, warning=FALSE}
work3

week <- c("1/10/2015", 
          "1/11/2015",
          "1/12/2015",
          "1/13.2015",
          "1/14/2015",
          "1/15/2015",
          "1/17/2015",
          "1/18/2015",
          "1/19/2015",
          "1/20/2015",
          "1/21/2015",
          "1/22/2015")

work3 |>
  filter(date %in% week, dept == "sewing") |>
  ggplot(aes(x = date, y=actual)) +
  geom_point()

sewing5 <- work3 |>
  filter(date %in% week, dept == "sewing") |>
  ungroup() |>
  select(-date, -day, -dept) 

head(sewing5)
```

### scatterplot matrix

```{r, message=FALSE, warning=FALSE}
sewing5 |>
  ggpairs()
```

### AIC variable selection

```{r}
n = nrow(sewing5)
mod0 = lm(actual ~ 1, data = sewing5)
mod.all = lm(actual ~., data = sewing5)
step(mod0, scope = list(lower = mod0, upper = mod.all))
```

```{r}
model.aic.new <- lm(actual ~ incentive + target, data = sewing5)
summary(model.aic.new)
```

```{r}
add1(model.aic.new, ~.+incentive*target, test = 'F')
```

p-value for F test for interaction is > 0.05 so we don't need interaction

### Check LINE assumptions

```{r}
library(broom)
model.table = augment(model.aic.new)
names(model.table)

ggplot(model.table, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, colour = 'blue') +
  labs(x = 'Fitted Values', y = 'Residuals') +
  ggtitle('Residual vs Fit') +
  theme_bw()

```

```{r}
ggplot(model.table, aes(sample = .resid)) +
  stat_qq() +
  stat_qq_line() +
  ggtitle('Normal Q-Q Plot') +
  theme_bw()
```

```{r}
shapiro.test(resid(model.3))

```

### transformation

```{r}

```









