---
title: "510 Project (Worker Productivity)"
output: html_document
date: "2022-12-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro to data set

Source:
https://www.kaggle.com/datasets/ishadss/productivity-prediction-of-garment-employees

Variables:
01 date : Date in MM-DD-YYYY
02 day : Day of the Week
03 quarter : A portion of the month. A month was divided into four quarters
04 department : Associated department with the instance
05 teamno : Associated team number with the instance 
06 noofworkers : Number of workers in each team 
07 noofstylechange : Number of changes in the style of a particular product
08 targetedproductivity : Targeted productivity set by the Authority for each team for each day. 
09 smv : Standard Minute Value, it is the allocated time for a task 
10 wip : Work in progress. Includes the number of unfinished items for products 
11 overtime : Represents the amount of overtime by each team in minutes
12 incentive : Represents the amount of financial incentive (in BDT) that enables or motivates a particular course of action.
13 idletime : The amount of time when the production was interrupted due to several reasons 
14 idlemen : The number of workers who were idle due to production interruption
15 actual_productivity : The actual % of productivity that was delivered by the workers. It ranges from 0-1.


# Some Questions. 

1. Which department is most productive?
2. Which day of the week is most productive for each department? 
4. Which team in the sewing department is the most productive?
5. Does the amount of over time impact the productivity in the sewing department? 
6. Does the amount of incentives impact the productivity in the sewing department? 
7. Does the number of workers impact the productivity in the sewing department? 

# Data cleaning

### 1. Read in the data set. Check if it's a tibble. 
```{r, message=FALSE, warning=FALSE}
library(readr)
work <- read_csv("~/Desktop/STAT510_F22/datasets/garments_worker_productivity.csv")
head(work)
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
is_tibble(work)
```


### 2. skim the data set. 

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
skim_without_charts(work)
```

There are a lot of missing values for wip. 

### 3. Investigate the missing values in wip

```{r, message=FALSE, warning=FALSE}
work |>
  mutate(is.na.wip = is.na(wip)) |>
  select(department, is.na.wip) |>
  group_by(department, is.na.wip) |>
  summarize(count = n())
```

The finishing department always has missing wip values because theytake all the "work in progress" parts and put them together. The sewing department will always have varying # of wip garments. 


### 4. select important variables and data cleaning

```{r}
work2 <- work |>
  select(-quarter, -smv, -idle_time, -idle_men, -no_of_style_change) |>
  mutate(department = factor(ifelse(department == "sweing", "sewing", department)))
names(work2)
```

```{r}
work3 <- work2 |>
  rename(dept = department, target = targeted_productivity, ot = over_time, 
         workers = no_of_workers, actual = actual_productivity) |>
  mutate(dept = factor(dept), team = factor(team), day = factor(day))
head(work3)
```

# Visualize distribution of variables of interest

```{r}
work3 |>
  ggplot( aes(x = dept)) +
  geom_bar()

work3 |>
  ggplot(aes(x = team)) + 
  geom_bar() + facet_wrap(~dept, nrow = 2)

work3 |>
  ggplot(aes(x = target)) + 
  geom_bar() + facet_wrap(~dept, nrow = 2)

work3 |>
  ggplot(aes(x = wip)) + 
  geom_histogram(bins = 100) + facet_wrap(~dept, nrow = 2)

work3 |>
  ggplot(aes(x = ot)) + 
  geom_histogram(bins = 50) + facet_wrap(~dept, nrow = 2)

work3 |>
  filter(dept == "sewing") |>
  ggplot(aes(x = incentive)) + 
  geom_histogram(bins = 100) + facet_wrap(~dept, nrow = 2)

work3 |>
  ggplot(aes(x = workers)) + 
  geom_histogram(bins = 50) + facet_wrap(~dept, nrow = 2)

work3 |>
  ggplot(aes(x = actual)) + 
  geom_histogram(bins = 50) + facet_wrap(~dept, nrow = 2)

```



# Visual Analysis 

### 1. Average actual productivity per department

```{r}
work3 |>
  group_by(dept) |>
  summarise(avg_productivity = mean(actual))
```

### 2. Average actual productivity per department and team

```{r}
library(ggplot2)
work.summary <- work3 |>
  group_by(dept, team) |>
  summarise(avg_actual= mean(actual)) |>
  ungroup()

work.summary

colors <- c("lightpink1", "lightpink2", "lightpink3", "mistyrose1", "mistyrose2", "mistyrose3", "rosybrown1", "rosybrown2", "rosybrown3", "thistle1", "thistle2", "thistle3")

work.summary |>
  ggplot(aes(x = team, y = avg_actual)) +
  geom_col(aes(fill = team)) +
  scale_fill_manual(values = colors) +
  facet_wrap(~dept, nrow = 2) + 
  theme_minimal() + theme(legend.position = "none")

```

